Toxic Code Snippets on Stack Overflow

### 2017-11-07 cloverflow-tse/cloverflow.tex
Checked occurences of violations.

\title{Toxic code snippets on Stack Overflow}

@@

% The paper headers
\markboth{IEEE TRANSACTIONS ON SOFTWARE ENGINEERING,~Vol.~X, No.~Y, Month~2017}%
{Ragkhitwetsagul \MakeLowercase{\textit{et al.}}: Outdated and License-Violating Code on Stack Overflow}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples in questions and answers. % They
Due to an absence of a checking mechanism after the code has been copied to
Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
being outdated or violating the original software license. We present a study
incorporating several state-of-the-art tools and techniques to automatically
extract and filter detected online clone pairs between 72,365 Java code snippets
on Stack Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
Our investigation of the online clone pairs revealed strong evidence that 154
clones have been copied from a Qualitas project to Stack Overflow.  We found 101
of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
we found 214 code snippets on Stack Overflow that violate the license of their
original software. We surveyed 201 high-reputation Stack Overflow answerers
(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
code. Only 3 answerers explicitly include software licenses in their answers and
138 answerers (69\%) never check for licensing conflicts between their copied code
snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

@@

Code cloning can also have side effects such as violating software
licenses or introducing software vulnerabilities. Carelessly cloning
code from one project to another project with a different license may
cause a software licensing violation~\cite{German2009}. This also
happens within the context of online Q\&A websites such as Stack
Overflow. An et al.~\cite{An2017} showed that 1,279 cloned snippets
between Android apps and Stack Overflow potentially violate software
licenses. Security is also among the main concerns when code is copied
from an online source. For example, Stack Overflow helps developers to solve
Android programming problems more quickly than other resources while,
at the same time, offers less secure code than books or the official
Android documentation~\cite{Acar2016}.

@@

We would like to motivate the readers by giving two examples of toxic
cod snippets.
The first example is outdated and license-violating online code clones in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{Hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part
of the accepted answer. The snippet shows how \textsf{Hadoop} implements the
{\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}}
class. The code snippet on the right shows another version of the same method,
but at this time extracted from the latest version (as of October 3, 2017) of
\textsf{Hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added on November
21, 2014. The added line is intended for cleaning up the reference in the
{\small{\verb|buffer|}} variable and avoid excess heap usage
(issue no.~HADOOP-11323\footnote{https://issues.apache.org/jira/browse/HADOOP-11323}).
While this change has already been introduced into the
{\small{\texttt{compare}}} method several years ago, the code example in Stack
Overflow post is still unchanged. 
%This example shows that inconsistencies
%between online code clones and their originals can lead users of Stack Overflow
%to reuse outdated code. 
In addition, the original code snippet of
{\small\texttt{WritableComparator}} class in Hadoop is distributed with Apache license
version 2.0 while its cloned instance on Stack Overflow contains only the
{\small{\texttt{compare}}} method and ignores its license statement on top of the
file. There are two potential issues for this. First, the code snippet may appear to be 
under Stack Overflow's CC BY-SA 3.0 instead of its original Apache
license. Second, if the code snippet is copied and incorporated into another
software project with a conflicting license, a legal issue may arise.

@@

We performed an empirical study of online code clones between Stack
Overflow and 111 Java open source projects to answer the following
research questions:
\begin{itemize}
	\item \textbf{RQ1 (Online code clones): }\textit{To what extent is source
		code cloned between Stack Overflow and open source projects?} We
	quantitatively measured the number of online code clones between Stack
	Overflow and open source projects to understand the scale of the
	problem. 
	\item \textbf{RQ2 (Patterns of online code clones): }\textit{Why do online
		code clones occur?} We categorised online clones into seven
	categories allowing insights to why online code clones are created.
	\item \textbf{RQ3 (Outdated online code clones): }\textit{Are
		online code clones up-to-date compared to their counterparts in the
		original projects?} We were interested in the outdated Stack
	Overflow code examples since users are potentially reusing
	them. 
	\item \textbf{RQ4 (Software licensing violation): }\textit{How often do
		licensing conflicts occur between Stack Overflow clones and their
		originals?} We investigated whether the reuse of online code clones
	can cause software developers to violate licenses.
	\item \textbf{RQ5 (Stack Overflow answerers' awareness): }

@@

We followed the 6 phases in the experimental framework
(\Cref{fig:exp_framework}) to answer our four research questions. To answer RQ1,
we rely on the number of manually validated true positive online clone pairs in
phase 3. We use the results of the manual classification by the seven patterns
of online code cloning to answer RQ2 (phase 4). For RQ3, we looked at the true
positive clone pairs that are classified as clones from Qualitas to Stack
Overflow and checked if they have been changed after cloning (phase 5).
Similarly, for RQ4, we looked at the license of each clone in the pattern QS,
EX, UD and checked for a possibility of licensing violation (phase 6). We use
the online survey to answer RQ5.

@@

\subsection{RQ4: Software Licensing Violation}
\vspace{0.25cm}
\textit{Do
	licensing conflicts occur between Stack Overflow clones and their
	originals?}
\vspace{0.25cm}

@@

\textbf{For RQ4, we found 214 code snippets on Stack Overflow that
	could violate the license of their original software. The majority of them
	do not contain licensing statements after they have been copied to
	Stack Overflow. For 164 of them, we were able to identify,
        with evidence, where the code snippet has been copied from.}

Note that the code snippets could violate the license, but do
not necessarily do so. In the example where the JFreeChart
developer copied his own code, she or he was free to change the
license. The same may have occurred with any of the 214 code
snippets.

@@

\textbf{Internal validity:} 
We applied
different mechanisms to ensure the validity of the clone
pairs we classified.  First, we used two widely-used
clone detection tools, Simian and SourcererCC.  We tried five other
clone detectors but could not add them to the study due to their
scalability issues and susceptibility to incomplete code snippets.
We adopted Bellon's agreement metric~\cite{Bellon2007}
to merge clone pairs for the manual
classification and avoid double counting of the same clone pairs. 
Nevertheless, our study might still suffer from false negatives, i.e.~online 
code clones that are not reported by the tools or are filtered 
out by the size (less than 10 lines) within the clone detection process.
% We also
% applied the Bellon's clone agreement to 4 different combinations of
% the tools' configurations.  For disagreed clone pairs, we applied
% several filters to remove spurious and uninteresting clones.

@@

The work that are closely similar to us are studies by An et al.~\cite{An2017}
and Abdalkareem et al.~\cite{Abdalkareem2017}. An et al.~investigated clones
between 399 Android apps and Stack Overflow posts. They found 1,226 code
snippets which were reused from 68 Android apps. They also observed that there
are 1,279 cases of potential license violations. The authors rely on the
timestamp to judge whether the code has been copied from/to Stack Overflow along
with confirmations from six developers. Instead of Android apps, we investigated
clones between Stack Overflow and 111 open source projects. Their results are
similar to our findings that there exist clones from software projects to Stack
Overflow with potential licensing violations. Abdalkareem et
al.~\cite{Abdalkareem2017} detected clones between Stack Overflow posts and
Android apps from the F-Droid repository and used timestamps to
determine the direction
of copying. They found 22 apps containing code cloned from Stack Overflow. They
reported that cloned code is commonly used for enhancing existing code. Their
analysis shows that the cloned code from Stack Overflow have negative effects to
quality of the apps. The median percentage of bug fixing commits after adding
Stack Overflow code (19.09\%) is higher than before adding the code (5.96\%)
with statistical significance.
%In our work, we defined seven patterns of online code cloning and 
%performed a large-scale manual check of 3,632 clone pairs. 
%We discovered 112 clone pairs with strong evidences, 
%based on natural text in comments and post contents, that they 
%were copied from Qualitas to Stack Overflow. By comparing the 
%clones to their latest versions in the software, we found that 
%64 code snippets on Stack Overflow are outdated and possibly harmful for reuse.

@@

We performed a detailed analysis of the online clone pairs on two
aspects: outdated code and licensing violation. The investigation of
the 154 clone pairs copied, with evidence, from Qualitas to Stack
Overflow reveals that 101 of them are outdated.  Moreover, we found 214
code snippets on Stack Overflow that violate the license of their
original software.

>>>>

\title{Toxic Code Snippets on Stack Overflow}

@@

% The paper headers
\markboth{IEEE TRANSACTIONS ON SOFTWARE ENGINEERING,~Vol.~X, No.~Y, Month~2017}%
{Ragkhitwetsagul \MakeLowercase{\textit{et al.}}: Toxic Code Snippets on Stack Overflow}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Online code clones are code fragments that are copied from software projects or
online sources to Stack Overflow as examples in questions and answers. % They
Due to an absence of a checking mechanism after the code has been copied to
Stack Overflow, they can become toxic code snippets, i.e.~they suffer from
being outdated or violating the original software license. We present a study
incorporating several state-of-the-art tools and techniques to automatically
extract and filter detected online clone pairs between 72,365 Java code snippets
on Stack Overflow and 111 Java open source projects in the curated Qualitas
corpus. We discovered and analysed 2,302 non-trivial online clone candidates.
Our investigation of the online clone pairs revealed strong evidence that 154
clones have been copied from a Qualitas project to Stack Overflow.  We found 101
of them (66\%) to be outdated and potentially harmful for reuse.  Furthermore,
we found 214 code snippets on Stack Overflow that could potentially violate the license of their
original software. We surveyed 201 high-reputation Stack Overflow answerers
(33\% response rate) and found that 131 of the answerers (65\%) have ever been notified of
outdated code in the answers and 26 of them (20\%) rarely or never fix the outdated
code. Only 3 answerers explicitly include software licenses in their answers and
138 answerers (69\%) never check for licensing conflicts between their copied code
snippets and Stack Overflow's CC BY-SA 3.0.
\end{abstract}

@@

Code cloning can also have side effects such as violating software
licenses or introducing software vulnerabilities. Carelessly cloning
code from one project to another project with a different license may
cause a software license violation~\cite{German2009}. This also
happens within the context of online Q\&A websites such as Stack
Overflow. An et al.~\cite{An2017} showed that 1,279 cloned snippets
between Android apps and Stack Overflow potentially violate software
licenses. Security is also among the main concerns when code is copied
from an online source. For example, Stack Overflow helps developers to solve
Android programming problems more quickly than other resources while,
at the same time, offers less secure code than books or the official
Android documentation~\cite{Acar2016}.

@@

We would like to motivate the readers by giving two examples of toxic
cod snippets.
The first example is an outdated and potentially license-violating online code clone in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{Hadoop}\footnote{\url{http://stackoverflow.com/questions/22315734}}.
\Cref{fig:before-after} shows---on the left---a code snippet embedded as a part
of the accepted answer. The snippet shows how \textsf{Hadoop} implements the
{\small{\texttt{compare}}} method in its {\small{\texttt{WritableComparator}}}
class. The code snippet on the right shows another version of the same method,
but at this time extracted from the latest version (as of October 3, 2017) of
\textsf{Hadoop}. We can see that they both are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added on November
21, 2014. The added line is intended for cleaning up the reference in the
{\small{\verb|buffer|}} variable and avoid excess heap usage
(issue no.~HADOOP-11323\footnote{https://issues.apache.org/jira/browse/HADOOP-11323}).
While this change has already been introduced into the
{\small{\texttt{compare}}} method several years ago, the code example in Stack
Overflow post is still unchanged. 
%This example shows that inconsistencies
%between online code clones and their originals can lead users of Stack Overflow
%to reuse outdated code. 
In addition, the original code snippet of
{\small\texttt{WritableComparator}} class in Hadoop is distributed with Apache license
version 2.0 while its cloned instance on Stack Overflow contains only the
{\small{\texttt{compare}}} method and ignores its license statement on top of the
file. There are two potential issues for this. First, the code snippet may appear to be 
under Stack Overflow's CC BY-SA 3.0 instead of its original Apache
license. Second, if the code snippet is copied and incorporated into another
software project with a conflicting license, a legal issue may arise.

@@

We performed an empirical study of online code clones between Stack
Overflow and 111 Java open source projects to answer the following
research questions:
\begin{itemize}
	\item \textbf{RQ1 (Online code clones): }\textit{To what extent is source
		code cloned between Stack Overflow and open source projects?} We
	quantitatively measured the number of online code clones between Stack
	Overflow and open source projects to understand the scale of the
	problem. 
	\item \textbf{RQ2 (Patterns of online code clones): }\textit{How do online
		code clones occur?} We categorised online clones into seven
	categories allowing insights to how online code clones are created.
	\item \textbf{RQ3 (Outdated online code clones): }\textit{Are
		online code clones up-to-date compared to their counterparts in the
		original projects?} We were interested in the outdated Stack
	Overflow code examples since users are potentially reusing
	them. 
	\item \textbf{RQ4 (Software licensing violation): }\textit{How often do
		license conflicts occur between Stack Overflow clones and their
		originals?} We investigated whether the reuse of online code clones
	can cause software developers to violate licenses.
	\item \textbf{RQ5 (Stack Overflow answerers' awareness): }

@@

We followed the 6 phases in the experimental framework
(\Cref{fig:exp_framework}) to answer our four research questions. To answer RQ1,
we rely on the number of manually validated true positive online clone pairs in
phase 3. We use the results of the manual classification by the seven patterns
of online code cloning to answer RQ2 (phase 4). For RQ3, we looked at the true
positive clone pairs that are classified as clones from Qualitas to Stack
Overflow and checked if they have been changed after cloning (phase 5).
Similarly, for RQ4, we looked at the license of each clone in the pattern QS,
EX, UD and checked for a possibility of license violation (phase 6). We use
the online survey to answer RQ5.

@@

\subsection{RQ4: Software Licensing Violations}
\vspace{0.25cm}
\textit{Do
	licensing conflicts occur between Stack Overflow clones and their
	originals?}
\vspace{0.25cm}

@@

\textbf{For RQ4, we found 214 code snippets on Stack Overflow that
	could potentially violate the license of their original software. The majority of them
	do not contain licensing statements after they have been copied to
	Stack Overflow. For 164 of them, we were able to identify,
        with evidence, where the code snippet has been copied from.}

Note that the code snippets could potentially violate the license, but do
not necessarily do so. In the example where the JFreeChart
developer copied his own code, she or he was free to change the
license. The same may have occurred with any of the 214 code
snippets.

@@

\textbf{Internal validity:} 
We applied different mechanisms to ensure the validity of the clone pairs we
classified.  First, we used two widely-used clone detection tools, Simian and
SourcererCC.  We tried five other clone detectors but could not add them to the
study due to their scalability issues and susceptibility to incomplete code
snippets. We adopted Bellon's agreement metric~\cite{Bellon2007} to merge clone
pairs for the manual classification and avoid double counting of the same clone
pairs. Nevertheless, our study might still suffer from false negatives,
i.e.~online code clones that are not reported by the tools or are filtered out
by the size (less than 10 lines) within the clone detection process. We selected
accepted answers on Stack Overflow in this study to focus on code snippets that
solve the question's problem and are often showed on top of the answer list.
While most of the time they are also the highest voted answers, some
times they are not. We plan to incorporate the highest voted answers in 
future work.
% We also
% applied the Bellon's clone agreement to 4 different combinations of
% the tools' configurations.  For disagreed clone pairs, we applied
% several filters to remove spurious and uninteresting clones.

@@

Work similar to ours are studies by An et al.~\cite{An2017}
Abdalkareem et al.~\cite{Abdalkareem2017}, and Baltes et al.~\cite{Baltes2017}. 
An et al.~investigated clones
between 399 Android apps and Stack Overflow posts. They found 1,226 code
snippets which were reused from 68 Android apps. They also observed that there
are 1,279 cases of potential license violations. The authors rely on the
timestamp to judge whether the code has been copied from/to Stack Overflow along
with confirmations from six developers. Instead of Android apps, we investigated
clones between Stack Overflow and 111 open source projects. Their results are
similar to our findings that there exist clones from software projects to Stack
Overflow with potential licensing violations. Abdalkareem et
al.~\cite{Abdalkareem2017} detected clones between Stack Overflow posts and
Android apps from the F-Droid repository and used timestamps to
determine the direction
of copying. They found 22 apps containing code cloned from Stack Overflow. They
reported that cloned code is commonly used for enhancing existing code. Their
analysis shows that the cloned code from Stack Overflow have negative effects to
quality of the apps. The median percentage of bug fixing commits after adding
Stack Overflow code (19.09\%) is higher than before adding the code (5.96\%)
with statistical significance.
Baltes et al.~\cite{Baltes2017} discovered that two
thirds of the clones from the 10 most frequently referenced Java code snippets on Stack
Overflow do not contain attributions.
%In our work, we defined seven patterns of online code cloning and 
%performed a large-scale manual check of 3,632 clone pairs. 
%We discovered 112 clone pairs with strong evidences, 
%based on natural text in comments and post contents, that they 
%were copied from Qualitas to Stack Overflow. By comparing the 
%clones to their latest versions in the software, we found that 
%64 code snippets on Stack Overflow are outdated and possibly harmful for reuse.

@@

We performed a detailed analysis of the online clone pairs on two
aspects: outdated code and potential license violation. The investigation of
the 154 clone pairs copied, with evidence, from Qualitas to Stack
Overflow reveals that 101 of them are outdated.  Moreover, we found 214
code snippets on Stack Overflow that could potentially violate the license of their
original software.
